---
title: "NREL ResStock 2025 — Data Orientation"
subtitle: "End-Use Load Profiles for U.S. Building Stock · resstock_amy2018_release_1"
---

## What is this dataset?

ResStock is NREL's large-scale simulation model of the U.S. residential building stock. The **2025 release** (`resstock_amy2018_release_1`) simulated roughly 550,000 representative homes using EnergyPlus with **2018 actual meteorological year (AMY)** weather — real observed conditions rather than the "typical" year used in most energy codes.

Each home was simulated under a **baseline (as-is)** condition and **32 upgrade scenarios** covering equipment replacements, envelope improvements, electric vehicles, and demand flexibility strategies. The result is a rich, nationally representative picture of residential energy use — how much, when, by what end use, and how it changes under different efficiency or electrification interventions.

**S3 root path:**

```
s3://oedi-data-lake/nrel-pds-building-stock/end-use-load-profiles-for-us-building-stock/2025/resstock_amy2018_release_1/
```

**OEDI browser:** [Open dataset in S3 viewer](https://data.openei.org/s3_viewer?bucket=oedi-data-lake&prefix=nrel-pds-building-stock%2Fend-use-load-profiles-for-us-building-stock%2F2025%2Fresstock_amy2018_release_1%2F)

---

## Folder Map

Files and folders at the root of `resstock_amy2018_release_1/`, with a one-line description of what each contains.

<pre class="filetree">
resstock_amy2018_release_1/
│   ~550K simulated U.S. homes · baseline + 32 upgrades · AMY 2018 weather
│
├── README_resstock_20251.pdf
│       Start here: overview, methods, and release notes
├── ResStockTechnicalReferenceGuide_2025_1.pdf
│       Full technical methodology and modeling assumptions
├── data_dictionary.tsv
│       Definitions for every column in the dataset
├── enumeration_dictionary.tsv
│       Lookup codes for all categorical variables
├── upgrades_lookup.json
│       Maps upgrade IDs 0–32 to scenario names
├── Column Name Crosswalk.csv
│       Column name mapping across dataset releases
│
├── metadata_and_annual_results/
│   │   Per-building characteristics + annual energy totals
│   ├── national/full/{csv,parquet}/
│   │       upgrade0 … upgrade32  (one file per scenario, all U.S.)
│   └── by_state/full/{csv,parquet}/
│           Same data, partitioned by state
│
├── timeseries_individual_buildings/
│   │   15-minute load profiles — one Parquet file per building
│   └── by_state/
│       └── upgrade={0–32}/state={XX}/
│               {building_id}-{upgrade}.parquet  (~2–3 MB each)
│
├── timeseries_aggregates/
│   │   Pre-summed load profiles by region (not per-building)
│   ├── by_state/
│   ├── by_ashrae_iecc_climate_zone_2004/
│   ├── by_building_america_climate_zone/
│   └── by_iso_rto_region/
│
├── weather/
│   │   AMY 2018 weather inputs used in EnergyPlus simulations
│   └── state={XX}/G[FIPS]_2018.csv
│           One CSV per county, standard EnergyPlus format
│
├── building_energy_models/
│   │   OpenStudio/EnergyPlus model input files
│   └── upgrade={0–32}/
│
├── cached_simulation_outputs/
│   │   Raw EnergyPlus simulation outputs
│   └── upgrade={0–32}/
│
├── geographic_information/
│   │   Spatial reference files for mapping and joining
│   ├── map_of_pumas_in_census_region_{1–4}_*.geojson
│   ├── map_of_us_states.geojson
│   └── spatial_tract_lookup_table.csv
│           Links census tracts to PUMA, state, climate zone …
│
└── comparisons/
        Side-by-side metrics vs. prior dataset releases
</pre>

---

## Folder Descriptions

### `metadata_and_annual_results/`

This is the most common starting point for analysis. It contains a **wide, flat table** where each row is one simulated building and columns cover two things: (1) the building's characteristics, and (2) its annual energy consumption.

**What the building characteristics columns include:**

- Location: state, PUMA (Public Use Microdata Area), county, climate zone
- Geometry: floor area, number of stories, foundation type, vintage (decade of construction)
- Envelope: insulation levels (attic, walls, foundation), window type, infiltration rate
- Systems: heating fuel and equipment type, cooling type, water heater type, duct location

**What the annual energy columns include:**

- Total site energy consumption (kWh, therms) broken out by end use: heating, cooling, water heating, lighting, refrigerator, dishwasher, clothes dryer, plug loads, etc.
- Both electricity and fuel-specific breakdowns

**Structure:**

```
metadata_and_annual_results/
├── national/full/{csv,parquet}/   → upgrade0 … upgrade32
└── by_state/full/{csv,parquet}/   → same data, state-partitioned
```

The `national/` files contain all ~550,000 buildings in a single file per upgrade scenario — convenient but large (Parquet files range from ~465 MB for the baseline to ~830 MB for some upgrades). The `by_state/` files hold the same data partitioned by state, which is useful for cloud query engines (Athena, DuckDB, Spark) that can push down state filters without reading the full table.

Both `csv/` and `parquet/` subdirectories are provided. Parquet is strongly preferred for performance; CSV is there if you need a format that opens without special tools.

Files are named `upgrade0.parquet`, `upgrade1.parquet`, ..., `upgrade32.parquet`. Upgrade 0 is the **baseline** (buildings as they currently exist); upgrades 1–32 represent hypothetical retrofits applied to the same set of buildings (see [Building Upgrades] below).

---

### `timeseries_individual_buildings/`

This folder contains the full sub-hourly timeseries for every simulated building — the most granular data in the dataset. Each Parquet file holds **one building's entire year of simulation output** at 15-minute intervals (~35,040 rows). The files are substantially richer than just energy consumption: they include indoor environmental conditions, emissions, airflow, occupant schedules, and weather.

**Structure:**

```
timeseries_individual_buildings/
└── by_state/
    └── upgrade={0–32}/
        └── state={XX}/
            └── {building_id}-{upgrade}.parquet
```

Files are named like `100102-0.parquet` — the number before the dash is the building's unique ID, the number after is the upgrade scenario. Individual files are small (~2–3 MB each), but the full dataset across all states and upgrades is enormous, so **always filter by state and upgrade** before downloading at scale.

**Column naming convention**

Columns follow a dot-delimited hierarchy: `out.{category}.{end_use}.{metric}..{unit}`. For example:

- `out.electricity.heating.energy_consumption..kwh`
- `out.natural_gas.hot_water.energy_consumption_intensity..kwh_per_ft2`
- `out.emissions.electricity.total.lrmer_midcase_25..co2e_kg`

The leading `out.` marks simulation outputs. `in.sqft` (floor area) is the one input characteristic carried into each file.

**What each file contains**

*Identifiers:* `bldg_id`, `timestamp`, `state`, `upgrade`, `in.sqft`

*Energy consumption* — absolute (kWh) and intensity (kWh/ft²) for each fuel × end use combination:

| Fuels | End uses |
|-------|----------|
| Electricity, Natural gas, Propane, Fuel oil | Heating, cooling, water heating, clothes dryer, clothes washer, dishwasher, refrigerator, freezer, lighting (interior, exterior, garage), plug loads, range/oven, ceiling fan, EV charging, pool/spa, well pump, PV generation, mechanical ventilation, heat pump backup |
| Site energy (all fuels combined) | Net and total |

```{r}
#| echo: false
#| message: false
library(reactable)
library(htmltools)

d <- data.frame(
  Category = c(
    "Emissions",
    "Thermal loads",
    "Indoor environment",
    "Outdoor conditions",
    "Airflow",
    "Hot water volumes",
    "Occupant schedules",
    "Unmet hours"
  ),
  Description = c(
    "CO₂e from electricity under multiple grid carbon scenarios (AER and LRMER variants at 15- and 25-year horizons). Single columns for natural gas, propane, and fuel oil combustion.",
    "Heating and cooling energy actually delivered to conditioned space — distinct from equipment consumption, which reflects efficiency losses. Includes hot water and heat pump backup loads.",
    "Dry bulb, operative, radiant, and dewpoint temperatures plus relative humidity and humidity ratio inside conditioned space, all at 15-minute resolution.",
    "Dry bulb, wet bulb, relative humidity, humidity ratio, diffuse solar radiation, direct normal solar radiation, and wind speed — echoed from the county-level AMY 2018 weather inputs.",
    "Infiltration, mechanical ventilation, and natural ventilation rates through the building envelope.",
    "Water volumes drawn by clothes washer, dishwasher, fixtures, and pipe distribution waste.",
    "Fractional multipliers (0–1) for appliances, lighting, occupancy, vacancy, and power outage events. Useful for understanding when simulated occupants are active.",
    "Hours per interval where the heating or cooling setpoint could not be maintained, and EV driving hours that could not be satisfied by available charge."
  ),
  Examples = c(
    "out.emissions.electricity.total.lrmer_midcase_25..co2e_kg\nout.emissions.natural_gas.total..co2e_kg",
    "out.load.heating.energy_delivered..kbtu\nout.load.cooling.energy_delivered..kbtu",
    "out.indoor_temperature.conditioned_space..c\nout.indoor_relative_humidity.conditioned_space..percentage",
    "out.outdoor_air_drybulb_temp..c\nout.weather.direct_normal_solar_radiation..watt_per_m2",
    "out.indoor_airflow.infiltration..cfm\nout.indoor_airflow.mechanical_ventilation..cfm",
    "out.hot_water.fixtures..gal\nout.hot_water.dishwasher..gal",
    "out.schedules.occupants\nout.schedules.lighting_interior",
    "out.unmet_hours.heating..hour\nout.unmet_hours.cooling..hour"
  ),
  Units = c(
    "kg CO₂e", "kBtu", "°C, %", "°C, W/m², m/s",
    "CFM", "gallons", "0–1 fraction", "hours"
  ),
  stringsAsFactors = FALSE
)

reactable(
  d,
  columns = list(
    Category = colDef(
      minWidth = 130,
      style = list(fontWeight = "600")
    ),
    Description = colDef(minWidth = 260),
    Examples = colDef(
      name = "Example columns",
      minWidth = 240,
      cell = function(value) {
        cols <- strsplit(value, "\n")[[1]]
        div(lapply(cols, function(x) tags$code(
          style = "display:block; font-size:0.76em; margin-bottom:3px;",
          x
        )))
      }
    ),
    Units = colDef(maxWidth = 115, align = "center")
  ),
  striped = FALSE,
  highlight = FALSE,
  bordered = TRUE,
  theme = reactableTheme(
    borderColor = "var(--rt-border-color)",
    headerStyle = list(fontWeight = "600")
  ),
  fullWidth = TRUE
)
```

This folder is the right place to go for **load shape analysis** — peak demand timing, load factor calculations, demand flexibility potential, indoor comfort analysis, or building-level profiles for grid studies. For annual totals or filtering buildings by characteristics, start with `metadata_and_annual_results/` instead.

---

### `weather/`

This folder contains the **actual meteorological year (AMY) 2018 weather data** used as inputs to the EnergyPlus simulations. AMY data reflects what actually happened in 2018 rather than a statistically "typical" year — meaning it captures the real heat waves, cold snaps, and cloud cover from that year.

**Structure:**

```
weather/
└── state={XX}/
    └── G[FIPS]_2018.csv   (one file per county)
```

Files are named using the county FIPS code — for example, `G0200130_2018.csv` is Alaska's Anchorage Borough (FIPS 02013). Each state folder contains one CSV per county in that state; Alaska alone has 29 files.

**What each file contains:**

The files are in standard EnergyPlus weather format with **hourly** data:

```{r}
#| echo: false
#| message: false
library(reactable)

w <- data.frame(
  Variable = c(
    "Dry bulb temperature",
    "Dew point temperature",
    "Relative humidity",
    "Atmospheric pressure",
    "Direct normal irradiance (DNI)",
    "Diffuse horizontal irradiance (DHI)",
    "Global horizontal irradiance (GHI)",
    "Wind speed",
    "Wind direction",
    "Precipitation"
  ),
  Description = c(
    "Air temperature measured in the shade, out of direct sunlight. The standard reference for outdoor temperature.",
    "Temperature at which air becomes saturated and condensation begins. A measure of absolute moisture content.",
    "Ratio of actual water vapor to the maximum possible at the current temperature.",
    "Weight of the air column above a location. Affects equipment performance in EnergyPlus calculations.",
    "Solar radiation arriving on a surface perpendicular to the sun's rays, from the sun's disc only.",
    "Solar radiation arriving on a horizontal surface from the sky (scattered, not direct).",
    "Total solar radiation on a horizontal surface — the sum of direct and diffuse components.",
    "Speed of horizontal air movement. Drives infiltration and exterior convection in the building model.",
    "Compass direction from which wind is blowing, in degrees (0° = North).",
    "Liquid water equivalent of precipitation. Used in moisture and envelope load calculations."
  ),
  Units = c("°C", "°C", "%", "Pa", "Wh/m²", "Wh/m²", "Wh/m²", "m/s", "degrees", "mm"),
  stringsAsFactors = FALSE
)

reactable(
  w,
  columns = list(
    Variable = colDef(
      minWidth = 175,
      style = list(fontWeight = "600")
    ),
    Description = colDef(minWidth = 320),
    Units = colDef(maxWidth = 90, align = "center")
  ),
  striped = FALSE,
  highlight = FALSE,
  bordered = TRUE,
  theme = reactableTheme(
    stripedColor = "#f8f9fa",
    highlightColor = "#e8f4fd",
    borderColor = "var(--rt-border-color)",
    headerStyle = list(fontWeight = "600")
  ),
  fullWidth = TRUE
)
```

To map a building to its weather file, use the county FIPS code in the `spatial_tract_lookup_table.csv` file in `geographic_information/`. The data dictionary identifies which metadata column holds each building's associated weather file location.

---

## Building Upgrades

The `upgrades_lookup.json` file maps upgrade IDs 0–32 to scenario names. Here's a plain-language breakdown of what each group represents.

**Upgrade 0 — Baseline**

No changes applied. Buildings are simulated as they currently exist, reflecting the real diversity of the U.S. housing stock — old and new, efficient and inefficient, gas-heated and electric. This is the reference point against which all other upgrades are compared.

---

**Upgrades 1–8 — Heating System Replacements**

These swap out the existing heating (and sometimes cooling) equipment for a more efficient alternative:

| ID | Scenario |
|----|----------|
| 1 | Natural Gas Furnace 95% AFUE (high-efficiency gas, ducted) |
| 2 | Propane or Fuel Oil Furnace 95% AFUE |
| 3 | Minimum-efficiency HVAC circa 2025 (code baseline) |
| 4 | Cold-climate ducted air source heat pump |
| 5 | Dual-fuel heating (heat pump + gas backup) |
| 6 | Single-speed geothermal heat pump |
| 7 | Dual-speed geothermal heat pump |
| 8 | Variable-speed geothermal heat pump |

Geothermal upgrades (6–8) use thermally enhanced grout and pipes for the ground loop. The varying speeds affect part-load efficiency and comfort.

---

**Upgrades 9–10 — Water Heater Replacements**

| ID | Scenario |
|----|----------|
| 9 | Heat pump water heater |
| 10 | High-efficiency natural gas tankless water heater |

---

**Upgrades 11–17 — Building Envelope Improvements**

These reduce heat loss and gain through the building shell. Some are individual measures; others combine measures that are typically done together:

| ID | Scenario |
|----|----------|
| 11 | Air sealing only |
| 12 | Attic floor insulation (unfinished attics) |
| 13 | Duct sealing and insulation |
| 14 | Drill-and-fill wall insulation + air sealing |
| 15 | Air sealing + attic insulation + duct sealing |
| 16 | Air sealing + attic insulation + duct sealing + wall insulation |
| 17 | Energy Star windows |

---

**Upgrade 18 — Deep Retrofit Package**

A bundled package: dual-speed geothermal heat pump + air sealing + attic insulation + duct sealing. Represents a comprehensive single-visit upgrade combining the most impactful envelope and equipment measures.

---

**Upgrades 19–23 — Electric Vehicle Adoption**

These add a household EV to the simulation, testing different charging assumptions:

| ID | Scenario |
|----|----------|
| 19 | EV with Level 1 charging (120V, ~1.4 kW) |
| 20 | EV with Level 2 charging (240V, ~7 kW) |
| 21 | Efficient EV with Level 2 charging |
| 22 | EV + Level 2 + demand flexibility (smart charging) |
| 23 | Efficient EV + Level 2 + demand flexibility |

EV scenarios are particularly useful for understanding load growth and peak demand implications of vehicle electrification.

---

**Upgrades 24–28 — HVAC Demand Flexibility**

These simulate smart thermostat strategies that shift heating/cooling load away from grid peak periods:

| ID | Strategy |
|----|----------|
| 24 | On-peak load shedding, 2°F setpoint offset |
| 25 | Pre-peak load shifting, 2°F offset, 4-hour pre-conditioning |
| 26 | On-peak load shedding, 4°F setpoint offset |
| 27 | Pre-peak load shifting, 4°F offset, 4-hour pre-conditioning |
| 28 | Pre-peak load shifting, 2°F offset, 1-hour pre-conditioning |

"Load shedding" means relaxing the thermostat setpoint during peak hours; "pre-peak load shifting" means pre-conditioning the home before the peak window begins. The offset (2°F vs. 4°F) controls how aggressively the thermostat deviates from the comfort setpoint.

---

**Upgrades 29–32 — Additional Packages**

| ID | Scenario |
|----|----------|
| 29 | General air sealing |
| 30 | General air sealing + attic insulation + duct sealing |
| 31 | Dual-speed geothermal heat pump + general air sealing + attic insulation + duct sealing |
| 32 | Variable-speed geothermal heat pump + general air sealing + attic insulation + duct sealing |

Upgrades 29–30 use a "general" air sealing specification (as opposed to the targeted approach in upgrade 11); upgrades 31–32 parallel the packages in upgrades 18 but with the general air sealing spec and different heat pump speeds.
