---
title: "Querying ResStock with DuckDB"
subtitle: "A walkthrough for accessing the ResStock S3 bucket"
---

# What's going on in here? 
This walkthrough covers how I would answer the following question:

> **1. How does indoor temperature relate to outdoor temperature in Philadelphia area single family attached homes?**

> **2. How much do energy do people consume for cooling during the summer?**

We'll use a tool called [`DuckDB`](https://duckdb.org/) to query both the [`annual_metadata`](https://data.openei.org/s3_viewer?bucket=oedi-data-lake&prefix=nrel-pds-building-stock%2Fend-use-load-profiles-for-us-building-stock%2F2025%2Fresstock_amy2018_release_1%2Fmetadata_and_annual_results%2F&limit=50) files and the individual building files from R. Ideally, this walkthrough will give you some tools that you can transfer to your question of interest. 

::: {.callout-caution}
This walkthrough assumes some knowledge of the [`{DBI}`](https://cran.r-project.org/web/packages/DBI/index.html) package and accessing databases in R. I'll touch on some of the concepts, but getting too in the weeds would make this post kind of long. For a more thorough introduction, see [Chapter 21 on databases from *R for Data Science 2e*](https://r4ds.hadley.nz/databases.html)

:::

# Setup


## What's duckdb?
Like any R analysis, the first thing we need to is install and load our package(s): 

```{r}
#| echo: true
#| code-fold: show
#| message: false 
#| warning: false

pacman::p_load('duckdb', 'tidyverse')
```

`{duckdb}` is a free, in-process database management system, which is a fancy way of storing and interacting with tabular data. It's smart, fast, and designed to work with hieararchical file systems like the ones 
stored in ResStock' S3 bucket. 

Using `{duckdb}` requires us write a little SQL, which is kind of an ugly, plain language way of interacting with data. But we'll get into that in a bit. 


::: {.callout-note}
All queries here run against OEDI's public S3 bucket, so no AWS credentials are needed. You do need an internet connection.
:::

## Establish database connection in R

When working in databases with R, the first thing that we need to do is establish a connection:

```{r}
#| echo: true
#| code-fold: show

temporary_con <- dbConnect(duckdb(), ":memory:") #' Temporary / default. Only lasts as long as r session
permanent_con <- dbConnect(duckdb(), 'data/db_storage.duckdb') #' A file path to store permanent results
```

First we tell R our database driver (duckdb), then we tell it how we want to store data. **For ResStock analyses, I recommend storing query results permanently, either locally or on a Drexel server.** As researchers, we often wind up with 10 dozen auxiallary questions, and it seems better to me to start analyses from a permanent place then hitting ReStock's s3 bucket every time. In the long run, this will also save you lots of time. 


## Loading s3 connection extentions for our database connection

The other thing we'll need is to load a few duckdb extensions that let it talk to websites. Like R, duckb has a package and extension ecosystem, which we can install and load like this:

```{r}
#| echo: true
#| code-fold: show
#| message: false
#| collapse: true

# Enable HTTP filesystem and S3 support
dbExecute(permanent_con, "INSTALL httpfs")
dbExecute(permanent_con, "LOAD httpfs")

# Public bucket — anonymous access
dbExecute(permanent_con, "SET s3_region = 'us-west-2'") #' Which sere
dbExecute(permanent_con, "SET s3_access_key_id = ''") #' Public, so no credentials needed
dbExecute(permanent_con, "SET s3_secret_access_key = ''") #' Public, so no credentials needed
```

In this step, I also configured our database to talk through data hosted on amazon's S3 server. 

And now we're setup!
---

# Establishing a connection

Remember I said that working with these things is like working with a file system that lives on the cloud? It took some setup, but we're finally able to make that real.

## Build a file path
Imagine that you were connecting to a file locally. You might do something like this:

```{r}
#| echo: true
#| code-fold: show
#| collapse: true

#' Path where data lives on S3 bucket
base_path <- 
  paste0("s3://oedi-data-lake/nrel-pds-building-stock/end-use-load-profiles-for-us-building-stock/2025/resstock_amy2018_release_1/")

base_path
```

You can think of that like your project root or working directory. 

Now, let's say that we want to start making some progress on our question. We know that we're interested in the Philadelphia CBSA, which comprises parts of Pennsylvania, Delaware and New Jersey. 

We can keep our footprint small by focusing on those states sub-directories. We'll start by building out some filepaths for upgrade 0 parquets. 

```{r}
#| echo: true
#| code-fold: show
#| collapse: true

states <-
  c('PA', 'DE', 'NJ')

mtd_ann_results <- glue::glue(
  "metadata_and_annual_results/by_state/full/parquet/state={states}/{states}_upgrade0.parquet"
)


st_file_paths <-
  paste0(base_path, mtd_ann_results)

st_file_paths
```

This gives us a 3 element character vector, where one vector element equals a state's file path. We can use these to (finally) connect to something. 

## Get annual table schema

Let's see what variables are available in the Pennsylvania table: 

```{r}
#| include: false
schema <-
  readr::read_csv('data/resstock_ann_table_schema.csv')
```
```{r}
#| echo: true
#| code-fold: show
#| eval: false 

schema <- dbGetQuery(permanent_con, sprintf(
  "DESCRIBE SELECT * 
   FROM read_parquet('%s') LIMIT 0",
  st_file_paths[1]
))

head(schema)

```

Now, we have a list of column names, formats, and some empty metadata that you don't really need to worry about. We can use this schema to look through and figure out what variables that we need to start answering our question. 


A few column naming conventions to know:

- `in.*` — building input characteristics (e.g., `in.state`, `in.heating_fuel`, `in.vintage`)
- `out.*` — simulation outputs (e.g., `out.electricity.total.energy_consumption`)
- `weight` — the statistical weight for each sample (how many real U.S. homes it represents)
- `bldg_id` — unique building identifier, links to individual timeseries files

Because, I've worked with this data a bit, I know that we can use these variables to identify which levels we'll need to filter for - both Philadelphia & single family attached. 

```{r}
#| echo: true
#| code-fold: show
#| collapse: true 



pa_house_cbsa_cats <-
dbGetQuery(permanent_con, 
  str_glue(
    "SELECT DISTINCT
        \"in.geometry_building_type_acs\", 
        \"in.metropolitan_and_micropolitan_statistical_area\",
     FROM read_parquet('{st_file_paths[1]}')"
  )
) 



level_names <- 
pa_house_cbsa_cats |> 
filter(
  str_detect(in.geometry_building_type_acs, 'Attached') & 
  str_detect(in.metropolitan_and_micropolitan_statistical_area, 'Philadelphia')
)

level_names


```

## Filter down to single family attached homes in Philly CBSA

Now, we want to get the identifiers associated with single family attached homes in Philly. Now we'll use our other paths, and some r iteration to make that happen.

```{r}
#| echo: true
#| code-fold: show
#| message: false
#| collapse: true

attached_ids <- map(
  st_file_paths,
  function(s3_path) {

    cli::cli_alert('trying {s3_path}:')
    dbGetQuery(
      permanent_con,
      str_glue(
        "SELECT bldg_id 
         FROM read_parquet('{s3_path}')
         WHERE \"in.geometry_building_type_acs\" = 'Single-Family Attached'
         AND \"in.metropolitan_and_micropolitan_statistical_area\" = 'Philadelphia-Camden-Wilmington, PA-NJ-DE-MD MSA'"
      )
    )
  }
) |> 
set_names(states)

map(attached_ids, nrow)
```

We have more ids then we know what to do with! (But more in PA, which makes sense given that's where Philly proper is).


# Individual Building Time Series

With the pre-selected `bldg_id` values, we can move to querying the individual building time series and looking at our outcomes of interst. The process will be the same, (specify path --> write query ---> get result), but this time, we'll use `dbExecute()` and save results to a permanent table.


## Specify Paths 

We can tack on the individual building time series path to our base path. 

```{r}
#| echo: true
#| code-fold: show
#| collapse: true

## Coerce to data frame to help stich together paths 
attached_ids_df <-
imap(attached_ids, ~.x |> mutate(state = .y)) |> 
list_rbind()

indvid_build_path <- 
  str_glue(
  'timeseries_individual_buildings/by_state/upgrade=0/state={attached_ids_df$state}/{attached_ids_df$bldg_id}-0.parquet')

build_path_fin <-
  paste0(base_path, indvid_build_path)

length(build_path_fin)

```


## Build the Query

Recall that building time series live in a slightly different place than the metadata that we just queried. These data are available in 15 increments for every building in the annual metadata for each day in 2018. This is a lot a lot of stuff, which is why we just spent time trying to focus our query. 

For our purposes, we don't really care about that 15-minute increments. Rather we'll just focus on daily values for the moment. This means that we need to write some sql to aggregate from minutes to days. 

```{r}
#| echo: true
#| code-fold: show
query_sfa_path <- paste0("'", build_path_fin, "'", collapse = ", ")


sfa_temps_query <-
  str_glue(
    "CREATE TABLE phl_sfa_temps AS 
SELECT

 ---- Time & Metadata Stuff 
  bldg_id,
  CAST(timestamp AS DATE)                        AS date,
  DAY(timestamp)                                 AS day,
  MONTH(timestamp)                               AS month,
  YEAR(timestamp)                                AS year,

---- Variables of Interest 
  AVG(\"out.indoor_operative_temperature.conditioned_space..c\") AS mean_op_indoor_temp,
  AVG(\"out.electricity.cooling.energy_consumption..kwh\") AS mean_cool_spend
FROM read_parquet([
    {query_sfa_path}
], hive_partitioning = true)
GROUP BY bldg_id, date, day, month, year
ORDER BY bldg_id, date;"
  )

if (!c('phl_sfa_temps') %in% dbListTables(permanent_con)) {
  dbExecute(permanent_con, sfa_temps_query)
} else {
  'Done, done, done'
}

## Did that work? 
tbl(permanent_con, 'phl_sfa_temps') |> head() |> collect()
```



---

## Adding outdoor weather

Same principles here. This time, we'll just point to a different part of the file structure. Remember that the weather file is stored in csvs, where each file represents a county. 


```{r}
#| echo: true
#| code-fold: true
#| message: false
#| collapse: true

#' Get Counties to build new file paths 
phl_counties <- map(
  st_file_paths,
  function(s3_path) {

    cli::cli_alert('trying {s3_path}:')
    dbGetQuery(
      permanent_con,
      str_glue(
        "SELECT DISTINCT \"in.county\"
         FROM read_parquet('{s3_path}')
         WHERE \"in.geometry_building_type_acs\" = 'Single-Family Attached'
         AND \"in.metropolitan_and_micropolitan_statistical_area\" = 'Philadelphia-Camden-Wilmington, PA-NJ-DE-MD MSA'"
      )
    )
  }
) |> 
set_names(states)

#' Get distinct counties to make things a little easier
phl_co <- imap(phl_counties, ~mutate(.x, st_abb = .y)) |> list_rbind()


#' build full file paths 
phl_weath_path <- str_glue('{base_path}weather/state={phl_co$st_abb}/{phl_co$in.county}_2018.csv')
phl_weath_path <- paste0("'", phl_weath_path, "'", collapse = ", ")

#' Build Query 
phl_weath_query <-

 str_glue("
CREATE TABLE phl_cbsa_co_weath AS
SELECT
  CAST(date_time AS DATE)                        AS date,
  MONTH(CAST(date_time AS DATE))                 AS month,
  YEAR(CAST(date_time AS DATE))                  AS year,
  regexp_extract(filename, '(G[0-9]+)_', 1)      AS county_id,

  -- drybulb
  AVG(\"Dry Bulb Temperature [°C]\")             AS mean_daily_drybulb,
  MIN(\"Dry Bulb Temperature [°C]\")             AS min_daily_drybulb,
  MAX(\"Dry Bulb Temperature [°C]\")             AS max_daily_drybulb,
  MAX(\"Dry Bulb Temperature [°C]\") - MIN(\"Dry Bulb Temperature [°C]\") AS range_daily_drybulb,

  -- relative humidity
  AVG(\"Relative Humidity [%]\")                 AS mean_daily_rh,
  MIN(\"Relative Humidity [%]\")                 AS min_daily_rh,
  MAX(\"Relative Humidity [%]\")                 AS max_daily_rh,
  MAX(\"Relative Humidity [%]\") - MIN(\"Relative Humidity [%]\") AS range_daily_rh,

  -- wind speed
  AVG(\"Wind Speed [m/s]\")                      AS mean_daily_wind_speed,
  MAX(\"Wind Speed [m/s]\")                      AS max_daily_wind_speed

FROM read_csv([{phl_weath_path}], hive_partitioning = true, filename = true)
GROUP BY date, month, year, county_id
ORDER BY county_id, date
")

#' Execute query
if (!c('phl_cbsa_co_weath') %in% dbListTables(permanent_con)){

dbExecute(permanent_con, phl_weath_query)} else{

  'Table Already made'
}


## Did that work? 
tbl(permanent_con, 'phl_cbsa_co_weath') |> 
head() |> 
collect()
```

## Last Random things

Just for fun, let's grab some additional housing characteristics so we can add a little more to our results.

```{r}
#| echo: true
#| code-fold: true 
#| collapse: true

phl_attached_characs <- map(
  st_file_paths,
  function(s3_path) {

    cli::cli_alert('trying {s3_path}:')
    dbGetQuery(
      permanent_con,
      str_glue(
        "SELECT  bldg_id, 
                \"in.county\",
                \"in.hvac_cooling_type\",
                \"in.vintage_acs\"
         FROM read_parquet('{s3_path}')
         WHERE \"in.geometry_building_type_acs\" = 'Single-Family Attached'
         AND \"in.metropolitan_and_micropolitan_statistical_area\" = 'Philadelphia-Camden-Wilmington, PA-NJ-DE-MD MSA'"
      )
    )
  }
) |> 
set_names(states) |> 
list_rbind()

```

---
# Visualizing Indoor & Outdoor Temps 

Let's calculate the **average indoor operative temperature** for our single family attached homes.

```{r}
#| echo: true
#| code-fold: show
#| warning: false
#| message: false

dbWriteTable(
  permanent_con,
  'phl_sfa_characs',
  phl_attached_characs,
  overwrite = TRUE,
  memory = TRUE
)

#' Average indoor operative time across all attached buildings,
#' By hvac & vintage
indoor_temp_tbl <- 
tbl(permanent_con, 'phl_sfa_temps') |>
 left_join(tbl(permanent_con, 'phl_sfa_characs'), by = join_by(bldg_id)) |> 
  summarise(
    daily_mean_indoor_temp = mean(mean_op_indoor_temp),
    .by = c('date', 'in.hvac_cooling_type', 'in.vintage_acs')
  ) |> 
  arrange(date) |> 
  collect()

#' Daily outdoor temp across all counties 
outdoor_weath_tbl <-
  tbl(permanent_con, 'phl_cbsa_co_weath') |> 
  summarise(
    mean_outdoor_temp = mean(mean_daily_drybulb),
    .by = c('date')
  ) |> 
  collect()

#' Join em up for the visual
temps_tbl_fin <-
  indoor_temp_tbl |> 
  left_join(outdoor_weath_tbl) |> 
  collect()
```

Anddd let's make a graph.

```{r}
#| fig-width: 8
#| fig-height: 7
#| echo: true
#| code-fold: true
#| outwidth: "100%"
theme_resstock <- function(base_size = 12, base_family = "") {
  theme_minimal(base_size = base_size, base_family = base_family) %+replace%
    theme(
      # Background
      plot.background = element_rect(fill = "#FFFFFF", color = NA),
      panel.background = element_rect(fill = "#F9F7FF", color = NA),

      # Grid
      panel.grid.major = element_line(color = "#DDD6FE", linewidth = 0.4),
      panel.grid.minor = element_blank(),

      # Axes
      axis.line = element_line(color = "#5B21B6", linewidth = 0.45),
      axis.ticks = element_line(color = "#5B21B6", linewidth = 0.35),
      axis.text = element_text(color = "#3D2A6B", size = rel(0.88)),
      axis.title = element_text(
        color = "#1E1B2E",
        size = rel(0.95),
        face = "bold",
        margin = margin(t = 6, r = 6)
      ),

      # Title / subtitle / caption
      plot.title = element_text(
        color = "#1E1B2E",
        size = rel(1.25),
        face = "bold",
        hjust = 0,
        margin = margin(b = 4)
      ),
      plot.subtitle = element_text(
        color = "#4A3F6B",
        size = rel(0.95),
        hjust = 0,
        margin = margin(b = 12)
      ),
      plot.caption = element_text(
        color = "#7C3AED",
        size = rel(0.78),
        hjust = 1,
        margin = margin(t = 10)
      ),
      plot.margin = margin(16, 20, 12, 16),

      # Legend
      legend.background = element_rect(fill = "#F9F7FF", color = NA),
      legend.key = element_rect(fill = "transparent", color = NA),
      legend.title = element_text(
        color = "#1E1B2E",
        face = "bold",
        size = rel(0.88)
      ),
      legend.text = element_text(color = "#3D2A6B", size = rel(0.85)),
      legend.position = "bottom",

      # Facet strips
      strip.background = element_rect(fill = "#EDE9FE", color = NA),
      strip.text = element_text(
        color = "#5B21B6",
        face = "bold",
        size = rel(0.88)
      )
    )
}

temps_tbl_fin |>
  filter(in.hvac_cooling_type != 'Non-Ducted Heat Pump') |>
  mutate(
    hvac_type = if_else(
      in.hvac_cooling_type %in% c('Central AC', 'Ducted Heat Pump'),
      'Ducted AC',
      in.hvac_cooling_type
    ),
    hvac_type = fct_relevel(hvac_type, "None", after = 0L)
  ) |>
  ggplot(aes(
    date,
    daily_mean_indoor_temp * 9 / 5 + 32,
    color = as.factor(in.vintage_acs)
  )) +
  geom_line() +
  geom_line(
    aes(date, mean_outdoor_temp * 9 / 5 + 32),
    color = 'black',
    alpha = .25,
    linetype = 'dashed'
  ) +
  facet_wrap(~hvac_type, nrow = 3, scales = 'free') +
  scale_color_viridis_d(option = "turbo", name = "Year-built") +
  labs(
    x = NULL,
    y = "Daily mean temperature (°F)",
    title = 'Indoor vs Outdoor Temperatures in Single Family Attached Homes',
    subtitle = 'A comparison of indoor and outdoor temperatures in 2018 | Philadelphi CBSA',
    caption = "Data from NREL ResStock's 2025 Output"
  ) +
  theme_resstock()
```


# Energy Consumption On Cooling 


```{r}
#| fig-width: 8
#| fig-height: 7
#| echo: true
#| code-fold: true
#| outwidth: "100%"
#| warning: false
#| message: false

indoor_spend_tbl <- 
tbl(permanent_con, 'phl_sfa_temps') |>
filter(month %in% c(6:9)) |> 
 left_join(tbl(permanent_con, 'phl_sfa_characs'), by = join_by(bldg_id)) |> 
  summarise(
    daily_mean_cool_spend = mean(mean_cool_spend),
    .by = c('date', 'in.hvac_cooling_type', 'in.vintage_acs')
  ) |> 
  arrange(date) |> 
  collect()

#' Daily outdoor temp across all co

#' Join em up for the visual
spend_tbl_fin <-
  indoor_spend_tbl |> 
  left_join(outdoor_weath_tbl) |> 
  collect()


spend_tbl_fin |>
  filter(!in.hvac_cooling_type %in% c('Non-Ducted Heat Pump', 'None')) |>
  mutate(
    hvac_type = if_else(
      in.hvac_cooling_type %in% c('Central AC', 'Ducted Heat Pump'),
      'Ducted AC',
      in.hvac_cooling_type
    ),
    hvac_type = fct_relevel(hvac_type, "None", after = 0L)
  ) |>
  ggplot(aes(
    date,
    daily_mean_cool_spend * 96,
    color = as.factor(in.vintage_acs)
  )) +
  geom_line() +
  # geom_line(
  #   aes(date, mean_outdoor_temp * 9 / 5 + 32),
  #   color = 'black',
  #   alpha = .25,
  #   linetype = 'dashed'
  # ) +
  facet_wrap(~hvac_type, nrow = 3, scales = 'free') +
  scale_color_viridis_d(option = "mako", name = "Year-built") +
  labs(
    x = NULL,
    y = "Daily total cooling energy (kWh)",
    title = 'Average Cooling Energy Consumption in Single Family Attached Homes',
    subtitle = 'A look at how much energy homes consume on cooling during the Summer | Philadelphia CBSA',
    caption = "Data from NREL ResStock's 2025 Output"
  ) +
  theme_resstock()

```





```{r}
#| echo: true
#| code-fold: true
tbl(permanent_con, 'phl_sfa_characs') |> summarise(n = n(), .by = c('in.hvac_cooling_type'))
```